---
title: "p8105_hw1_cjd2195"
output: github_document
date: "2023-09-21"
---

## Problem 1

First up: let's load our moderndive library and load the "early_january_weather" dataset. I will also go ahead and look at the help file for this dataset to learn more about its components and structure.

```{r}
library(moderndive)
data("early_january_weather")
?early_january_weather
```

This dataset has 15 different variables! They are: 

* origin: where the measurement was taken
* year: year the measurement was taken
* month: month the measurement was taken
* day: day the measurement was taken
* hour: hour the measurement was taken
* temp: temperature in F
* dewp: dewpoint in F
* humid: relative humidity
* wind_dir: wind direction, in degrees
* wind_speed: wind speed, in mph
* wind_gust: wind gust speed, in mph
* precip: precipitation, in inches
* pressure: sea level pressure, in millibars
* visib: visibility, in miles
* time_hour: date and hour of the recording as a `POSIXct` date

To find out how big our dataset is, we could look at the help file, but I'll also call some R functions to help me figure it out. 

```{r, eval = FALSE}
nrow(early_january_weather)
ncol(early_january_weather)
```

There are `r nrow(early_january_weather)` rows in the dataset, and `r ncol(early_january_weather)` columns in the dataset. 

We can also do some simple investigative analyses with these data, like assessing the mean recorded temperature. In this case, we can check the mean temperature. First I'll save the data to a dataframe, which means I'll also need to load the tidyverse library.  

```{r, message = FALSE}
library(tidyverse)

early_jan_weather_df = tibble(
  early_january_weather
)

mean_temp = mean(pull(early_jan_weather_df, temp))
```

The mean temperature is `r mean_temp` degrees F.

Lastly, let's check out a quick visualization of our temperature data over time to see if we can observe anything. 

```{r}
ggplot(early_jan_weather_df, aes(x = time_hour, y = temp, color = humid)) + geom_point()
ggsave("temp_vs_time_scatter.png")
```

This plot appears to show some sort of cyclic pattern in which temperatures rise and fall over short intervals (probably daytime vs. nighttime temperatures), but gradually over the span of all data collected, temperatures were on the rise. The minimum temperature ovserved was below 25degrees, and the maximum was above 55degrees. Across the data, we see mostly drier measurements with respect to humidity, but there are some stretches where humidity appears to raise to close to 100% (likely indicating some sort of precipitation). 

## Problem 2

We are going to make a dataframe with several specified variable types: 

* A random sample, n = 10, from a standard normal distribution
* A logical vector that will tell us if those samples are greater than 0
* A character vector, len = 10
* A factor vector, len = 10, with 3 different levels

```{r}
factor_levels = c("red", "blue", "green")
problem_2_df = tibble(
  random_sample_var = rnorm(10),
  logical_var = random_sample_var > 0,
  character_var = c("This", "is", "an", "example", "of", "a", "character", "vector", "length", "ten"),
  factor_var = factor(c("red", "blue", "green", "green", "blue", "red", "red", "blue", "blue", "blue"), levels = factor_levels)
)
```

Let's take a peek at what trying to call the mean value for each variable yields.

For the `random_sample_var`:
```{r}
mean(pull(problem_2_df, random_sample_var))
```

For the `logical_var`:
```{r}
mean(pull(problem_2_df, logical_var))
```

For the `character_var`:
```{r}
mean(pull(problem_2_df, character_var))
```

For the `factor_var`:
```{r}
mean(pull(problem_2_df, factor_var))
```

Taking the mean works for the numeric `random_sample_var`, and the logical `logical_var`, but not for the character and factor variables. 

Let's see what happens when we try to force the variables to be numeric ones. 

```{r, eval = FALSE}
as.numeric(pull(problem_2_df, random_sample_var))
as.numeric(pull(problem_2_df, logical_var))
as.numeric(pull(problem_2_df, character_var))
as.numeric(pull(problem_2_df, factor_var))
```

The conversion for the `random_sample_var` completes without any issues, because it's already a numeric variable.

The conversion of the `logical_var` also does not yield an error, but instead converts `FALSE` to `0` and `TRUE` to `1`, which helps explain why we have a mean value of 0.5 for this variable.

With the `character_var`, every value becomes `NA` with the warning that they have been introduced by coercion. This is consistent with the `NA` that was returned when we tried to take the mean above, and the warning explicitly states that non-numeric or logical variables cannot have a mean.

Finally, the `factor_var` conversion completes by assigning a number to each of the levels represented in the vector (a total of 3 in this case). Despite this successul conversion to numeric, when we tried to take the mean previously we still yielded the warning about non-numeric or logical variables and returned `NA`. This warning explicitly explains why we wouldn't get a value for the mean like we did with the logical vector; even though both were converted by as.numeric, fundamentally, the logical_vector can have a mean while the factor cannot.
